### 2021-05-26 LaTeX-ACM/sections/appendix.tex
edits + resolving merge conflicts

\section{Introductory Remarks}

For better or worse, blockchain technologies like Ethereum have dramatically lowered the barrier to entry for developing and deploying financial technology. New tokens have been launched with a few clicks of a user interface, and large investment infrastructures have been developed and deployed with little regulatory oversight. Blockchain exchange services allow order-based trading of digital currencies, tokens, and other digital assets. Such exchanges are a key component to blockchain-based economic activity.

Financial regulators seek to provide consumer protection during the issuing and trading of financial products and assets. They are concerned by their limited ability to intervene when trading is conducted on decentralized networks that, like Ethereum, run on the open internet. Order-driven trading can happen fully `on-chain' and this was experimented within the early days of Ethereum, but it has been largely abandoned for performance reasons in favour of running on centralized servers. More specifically, the core functionality is performed off-chain (\eg matching orders) while other aspects (\eg loading accounts, order cancellation) might be performed on-chain. In this world, company names, employees, addresses, and publicly addressable servers all provide regulatory hooks.

Our research group is actively collaborating with our jurisdiction's (Quebec, Canada) financial regulator, the Autorité des Marchés Financiers (AMF), to help them forecast how trading can be impacted by blockchain technology. This paper addresses their concerns about the feasibility of a `worst-case scenario' where a trading platform is anonymously deployed on a public blockchain, like Ethereum, and runs autonomously without any further intervention from an externally visible entity. Such a design appears feasible but is considered too slow. Together, we agreed it is a very good time to do a deep dive into understanding precisely \textit{how slow} for the following reasons: (1) public blockchains are becoming faster (both in theory and in practice) providing future efficiency gains for on-chain trading, (2) demand for on-chain trading is exemplified by the recent popularity of dealer quote-based trading like Uniswap and Curve Finance (reviewed below), and (3) stablecoins have become popular and allow on-chain trading with pricing in USD, alleviating another regulatory hook: the need for platforms to maintain traditional accounts for holding governmental currency and interfacing with the banking system.

\paragraph{Contributions.} We study fully on-chain markets through design and implementation with Solidity---a high-level programming language for Ethereum that is syntactically similar to Java~\cite{Ethereum41:online}. We choose Ethereum as a hostile environment for an order book: Ethereum is fully decentralized (hardest to regulate), network participation is open to anyone on the internet (strongest adversarial model), and, as a result, it is slow (a lower-bound benchmark). Generally if an application is feasible on Ethereum, it will also be feasible and only run faster on a private (or permissioned) blockchain (\eg Hyperledger), which is a blockchain operated by authorized network nodes only, generally visible to regulators.

Our proof-of-concept, \cm, is an extensible base class suitable for experiments. \cm is designed from a security perspective: users only trust their assets to \cm's auditable code (non-custodial) and not to a third party; all operations are transparent; and we nearly eliminate the ability for adversarial network nodes to profit from front-running orders. In fact, we believe we are the first to design a decentralized application that actually leverages front-running (\ie miner extractable value~\cite{daian2019flash}) for the benefit of the system (see Section~\ref{sec:close}).

While Solidity and its compiled bytecode is like many common programming languages, it also has quirks that require experimentation to best optimize performance (\eg factoring in the gas costs of operations, gas refunds, limits to Solidity's object oriented design, clearing mappings, \etc). We test five priority queues---the core data structure of the call market---and various options for cleaning up our data once finished with it. The bottom line is that the current benchmark for a \cm-esque design is in the low hundreds of trade executions per block on Ethereum today. This positions \cm as a feasible design for only a narrow set of markets today (low liquidity, small number of traders) which is good news for regulators. However we caution that the technology is likely to improve vastly in the coming years and we demonstrate one avenue for improvement through `roll-ups'~\cite{kalodner2018arbitrum}. While roll-ups use centralized servers and are not fully on-chain, an architecture where these servers only interface with Ethereum and never interface directly with users is still concerning, as it side-steps the regulatory hook that is present in many other server-augmented trading platforms.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\section{Preliminaries}

\subsection{Ethereum and Blockchain Technology}

A public blockchain is an open peer-to-peer network that maintains a set of transactions without a single entity in charge. In Ethereum, \emph{transactions} encode the bytecode of user-written \emph{decentralized applications} (DApps) to be stored on the blockchain; and the function calls made to the DApp. Every execution of every function call is a validated by all honest, participating nodes to correct; a property that is robust against a fraction of faulty and malicious network nodes (or more precisely, their accumulated computational power). Once transactions are agreed upon, all honest participants will have identical sets of transactions in the same order. For Ethereum, this is conceptualized as the current state of a large \emph{virtual machine (EVM)} that is running many DApps.

Transactions are broadcast by users to the blockchain network where they are propagated to all nodes. Nodes that choose to \emph{mine} will collect transactions (in the order of their choosing) into a block, and will attempt to have the network reach a consensus that their block should be added to the set (or chain) of previous blocks. A transaction is considered finalized once consensus on its inclusion has held for several additional blocks.

\textbf{Ethereum's Gas Model.} Every transaction results in the participating nodes having to execute bytecode. This is not free. When a transaction is executed, each opcode in the execution path accrues a fixed, pre-specified amount of \emph{gas}. The function caller will pledge to pay a certain amount of Ethereum's internal currency \emph{ETH} (typically quoted in units of Gwei which is one billionth of an ETH) per unit of gas, and miners are free to choose to execute that transaction or ignore it. The function caller is charged for exactly what the transaction costs to execute, and they cap the maximum they are willing to be charged (\textit{gas limit}). If the cap is too low to complete the execution, the miner keeps the Gwei and \emph{reverts} the state of the EVM (as if the function never ran).

A miner can include as many transactions (typically preferring transactions that bid the highest for gas) that can fit under a pre-specified \textit{block gas limit}, which is algorithmically adjusted for every block. As of the time of writing, the limit is approximately 11M gas. Essentially, our main research question is how many on-chain trades can be executed without exceeding that limit. Later we also discuss several bytecode operations (\emph{opcodes}) that refund gas (\ie cost negative gas), which we heavily utilize in our optimizations.

\subsection{Trade Execution Systems}

 There are three main approaches to arranging a trade~\cite{Har03}. In a \emph{quote-driven} market, a dealer uses its own inventory to offer a price for buying or selling an asset. In a \emph{brokered exchange}, a broker finds a buyer and seller. In an \emph{order-driven} market, offers to buy (\emph{bids}) and sell (\emph{offers}/\emph{asks}) from many traders are placed as orders in an order book. Order-driven markets can be \emph{continuous}, with buyers/sellers at any time adding orders to the order book (\emph{makers}) or executing against an existing order (\emph{takers}); or they can be \emph{called}, where all traders submit orders within a window of time and orders are matched in a batch (like an auction).

\subsection{Trading Blockchain-based Assets}

%\input{tables/eval.tex}

\paragraph{Central Exchanges (CEX).} Traditional financial markets (\eg NYSE and NASDAQ) use order-matching systems to arrange trades. An exchange will list one or more assets (stocks, bonds, derivatives, or more exotic securities) to be traded with each other, given its own order book priced in a currency (\eg USD). Exchanges for blockchain-based assets (also called crypto assets by enthusiasts) can operate the same way, using a centralized exchange (CEX) design where a firm (\eg Binance, Bitfinex, \etc) operates the platform as a trusted third party in every aspect: custodianship over assets/currency being traded, exchanging assets fairly, offering the best possible price execution. Security breaches and fraud in centralized exchanges  (\eg MtGox~\cite{TheHisto45:online}, QuadrigaCX~\cite{SEBIOrde83:online}, and many others) have become a common source of lost funds for users, while accusations of unfair trade execution have been levelled but are difficult to prove. Today, CEXes are often regulated as other money service businesses---this provides some ability for the government to conduct financial tracking but does little to provide consumer protection against fraud.

\paragraph{On-chain Order Books.} For trades between two blockchain-based assets (\eg a digital asset priced with a cryptocurrency, stablecoin, or second digital asset), order matching can be performed `on-chain' by deploying the order-matching system either on a dedicated blockchain or inside a decentralized application (DApp; \textit{a.k.a.} smart contract). In this model, traders entrust their assets to an autonomously operating DApp with known source code instead of a third party custodian that can abscond with or lose the funds. The trading rules will operate as coded, clearing and settling can be guaranteed, and order submission is handled by the blockchain---a reasonably fair and transparent system (but see front-running below). Finally, anyone can create an on-chain order book for any asset (on the same chain) at any time. While they sound ideal, performance is a substantial issue and the main subject of this paper. Since it is an open system, there is no obvious regulatory hook (beyond the blockchain itself).

In this paper, we focus on benchmarking an order book for the public blockchain Ethereum. Ethereum is widely-used and we stand to learn the most from working in a performance-hostile environment. Exchanges could be given their own dedicated blockchain, where trade execution logic can be coded into the network protocol. Trading systems on permissioned blockchains (\eg NASDAQ Linq, tZero) can also improve execution time and throughput, but they reduce user transparency and trust if unregulated.

\paragraph{On-chain Dealers.} An advantage of on-chain trading is that other smart contracts, not just human users, can initiate trades, enabling broader decentralized finance (DeFi) applications. This has fuelled a resurgence in on-chain exchange but through a quote-driven design rather than an order-driven one. Automated market makers  (\eg Uniswap) have all the trust advantages of an on-chain order book, plus they are very efficient relative to an on-chain order book. The trade-off is that they operate as a dealer---the DApp exchanges assets from its own inventory. This inventory is loaded into the DApp by an investor who will not profit from the trades themselves but hopes their losses (termed `impermanent losses') are offset over the long-term by trading fees. By contrast, an order book requires no upfront inventory and trading fees are optional. Finally, there is a complicated difference in their price dynamics (\eg market impact of a trade, slippage between the best bid/ask and actual average execution price, \etc)---deserving of an entire research paper to precisely define. We leave it as an assertion that with equal liquidity, order books have more favourable price dynamics for traders.

\paragraph{Hybrid Designs.} Before on-chain dealers became prominent in the late 2010s, the most popular design was hybrid order-driven exchanges with some trusted off-chain components and some on-chain functionality. Such decentralized exchanges (DEXes) were envisioned as operating fully on-chain, but performance limitations drove developers to move key components, such as the order matching system, off-chain to a centralized database. A landscape of DEX designs exist (\eg EtherDelta, 0x, IDEX, \etc): many avoid taking custodianship of assets off-chain, and virtually all (for order-driven markets) operate the order book itself off-chain (a regulatory hook). A non-custodial DEX solves the big issue of a CEX---the operator stealing the funds---however trade execution is still not provably fair, funds can still be indirectly stolen by a malicious exchange executing unauthorized trades, and server downtime is a common frustration for traders. An enhancement is to prove that trade execution is correct (\eg Loopring) but these proofs have blindspots (discussed in section~\ref{sec:rollups}).

\subsection{Related Works}

\paragraph{Blockchain Limitations and Solution.} While an order book is a ledger and blockchains provide a distributed ledger, it is not straightforward to drop a continuous-time order book onto a blockchain. An older 2014 paper~\cite{clark2014decentralizing} on the `Princeton prediction market'~\cite{Bra13} motivates our work. The authors observe the following limitations of on-chain continuous order books: block intervals are slow and not continuous, there is no support for accurate time-stamping, transactions can be dropped or reordered by miners, and fast traders can react to submitted orders/cancellations when broadcast to network but not in a block and have their orders appear first (as examined in later work on front-running: \cite{eskandari2019sok,daian2019flash}).

\paragraph{Call Markets.} The researchers propose using a call market instead of a continuous-time market~\cite{clark2014decentralizing}. Orders are collected and placed into the order book over a window of time (\eg 1 or more blocks), then the market is closed and the orders are processed in batch: the best bids are matched to the best asks in order. If the prices overlap, the miner keeps the difference (which they could extract anyways through front-running). Call markets largely side-step front-running attacks from other traders because reordering trades has no impact (discussed more in section~\ref{sec:front}). The paper does not include an implementation and was envisioned as running on a custom blockchain (Ethereum was still in development in 2014). Market operations are part of the blockchain logic.

Large exchanges, like the NYSE and NASDAQ, run two call markets every trading day in parallel with a continuous-time market. One call market closes at market open to produce the opening price for each stock, while the other closes at the end of the day to produce the closing price. Other exchanges, called crossing networks, also operate as a call market at various times throughout the trading day.\footnote{A crossing network uses a secondary market for determining the closing price. Many prominent crossing networks are operated internally within a brokerage for its clients, and often as a `dark pool' with an unpublished order book.} Call markets are studied widely in finance~\cite{Har03}. Time-sensitive traders submit orders early, especially in crossing networks that close at a randomly determined time (traders risk missing the call if they wait too long). A blockchain happens to provide this function naturally, as blocks are published unpredictably. Price-sensitive traders wait to base their pricing off the already submitted orders and do not mind missing a call if it obtains them a better price.

\paragraph{Other Academic Literature.} There are numerous industry projects on blockchain-based exchanges and order books but most of the academic literature is on topics that are related but tangential to the mechanics of trade execution. Early (and some recent) literature consider trade execution under encryption (\ie dark markets) for securities~\cite{TP07,YSLT10,TW12,cartlidge2019mpc} and futures~\cite{massacci2018futuresmex}. A number of projects consider structuring derivatives in smart contracts---Velocity ~\cite{eskandari2017feasibility}, Findel ~\cite{biryukov2017findel}---but once issued, the derivative can be traded using exchanges. Atomic swaps (\ie payment vs delivery) are necessary for settling trades and some general approaches include Arwen~\cite{heilman2020arwen} and Tesseract~\cite{bentov2017tesseract}.

The most similar academic work to this paper is the TEX exchange~\cite{khalil2019tex}. TEX is also an implementation of an order book, however it uses on-chain and off-chain components, and does not answer our research question of how feasible an on-chain order book is. We later compare it to our variant with roll-ups in Section~\ref{sec:rollups}. The main overlap is our common focus on resisting front-running attacks. In a recent survey on front-running attacks in blockchain~\cite{eskandari2019sok}, three solutions are proposed. One solution is to provide confidentiality over the content of orders which is the approach taken by TEX. The main downside is that honest traders cannot submit their orders and leave, they must interact in a second round to reveal their orders. The second solution is to sequence transactions according to some rule, either at the protocol level~\cite{MZGA20} or as a third party service~\cite{Kla20} (a regulatory hook). These are very recent works and are not available for experimentation on Ethereum yet (although Chainlink has announced an intention\footnote{A. Juels. \href{https://blog.chain.link/chainlink-fair-sequencing-services-enabling-a-provably-fair-defi-ecosystem/}{blog.chain.link}, 11 Sep 2020.}). The third solution is to design the service in a way that front-running attacks are not profitable---this is the approach we take here.

% https://eprint.iacr.org/2020/269.pdf
% Wendy, the Good Little Fairness Widget – Achieving Order Fairness for Blockchains

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\section{Priority Queues}\label{sec:pq}

\input{tables/pq.tex}

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, closing the market and processing all the orders is the most time consuming step. The most critical design decision is the data structure for holding orders. While data structures are well studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object oriented programming) that create difficulties in assessing which will perform best without actually deploying and evaluating each variant.

When closing a call market, the orders are examined in order: highest to lowest price for bids, and lowest to highest price for asks. In most circumstances, the market closing algorithm does not have to consider any deeper bids/asks from the list when choosing whether the current best bid and ask can be fulfilled. The only exceptions are in the case of a tie on price or a cancelled order, both of which we return to later. For this reason, the ideal data structure for storing bids/asks is a \textit{priority queue} (see Table~\ref{tab:PQ_API}) where each order's priority is its price. Specifically, we use two PQs---one for bids where the highest price is the highest priority, and one for asks where the lowest price is the highest priority.

There are numerous ways of implementing a PQ. A PQ has an underlying list---common options include a static array, dynamic array, and linked list. The most expensive operation is keeping the data sorted---common options include (i) sorting during each enqueue, (ii) sorting for each dequeue, or (iii) splitting the difference by using a heap as the underlying data structure. Respectively, the time complexities are (i) linear enqueue and constant dequeue, (ii) constant enqueue and linear dequeue, and (iii) logarithmic enqueue and logarithmic dequeue. As closing the market is very expensive with any PQ, we rule out using (ii) as fully sorting while dequeuing would be prohibitive. We experiment with the following 5 options for (i) and (iii):

\item \textbf{Heap with Dynamic Array.} A heap is a type of binary tree data structure that comes in two forms of a (i) Max-Heap and (ii) Min-Heap. All the nodes of the tree are in a specific order and the root always represents the highest priority item of the data structure (the largest and smallest values in the Max-Heap and Min-Heap respectively). We implement a PQ with a heap that stores its data in a dynamically-sized array.

\item \textbf{Heap with Static Array.} A heap can be also represented by a Solidity storage array in which the storage is statically allocated. To do this, we pass the required size of the array as a constructor parameter to the PQ smart contract.

\item \textbf{Heap with Mapping.} In the above implementations, the entire order is stored (as a struct) in the heap. In this variant, we store the order struct in a Solidity mapping and store the mapping keys in the heap.

\item \textbf{Linked List.} In this variant, we insert a new element into its correct position (based on its price) when running enqueue. The PQ itself stores elements in a linked list (enabling us to efficiently insert a new element between two existing elements). Solidity is described as object-oriented but the equivalent of an object is an entire smart contract. Therefore an object-oriented linked list must either (i) create each node in the list as a struct---but this is not possible as Solidity does not support recursive structs---or (ii) make every node in the list its own contract. The latter option seems wasteful and unusual, but we try it out anyways. Thus each node is its own contract and contains the order data and a pointer to the address of the next contract in the list.

\item \textbf{Linked List with Mapping.} Finally, we try a variant of a linked list using a Solidity mapping. The value of the mapping is a struct with order data and the key of the next (and previous) node in the list. The contract stores the key of the first node (head) and last node (tail) in the list.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\subsection{Priority Queue Evaluation}

\paragraph{Enqueue Performance.} We implemented, deployed, and tested each PQ using Truffle and Ganache. We tried a variety of tests (including testing the full call market with each variant) with consistent results in performance. A simple test to showcase the performance profile is shown in Figure~\ref{fig:random_insertion}. We simply enqueue 50 integers chosen at random from a fixed interval in each PQ variant. The bigger the PQ gets, the longer enqueue takes---a linear increase for the linked list variants, and logarithmic for the heap variants.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{fig/random_insertion3.png}
\caption{Gas costs for enqueuing 50 random integers into five priority queue variants. For the x-axis, a value of 9 indicates it is the 9th integer entered in the priority queue. The y-axis is the cost of enqueuing in gas.  \label{fig:random_insertion}}
\end{figure}

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

\input{tables/dequeue.tex}

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

\paragraph{Dequeue Performance.} For each PQ variant storing 50 random integers, the \texttt{Dequeue()} function is iterated until the data structure is empty. The total gas cost for fully dequeuing the PQ variants is outlined in Table~\ref{tab:PQUnitTests}. These tests are performed using the following Ethereum gas metrics: block gas limit $=$11,741,495 and 1 gas $=$ 56 Gwei.\footnote{EthStats (July 2020): \url{https://ethstats.net/}} Dequeuing removes data from the contract's storage. Recall this results in a gas refund. Based on our manual estimates (EVM does not expose the refund counter), every variant receives the maximum gas refund possible (\ie half the total cost of the transaction). In other words, each of them actually consumes twice the \texttt{gasUsed} amount in gas before the refund, however none of them are better or worse based on how much of a refund they generate.

\paragraph{Discussion.} Based on enqueuing, the heap variants are the cheapest in terms of gas, while based on dequeuing, the link list variants are the cheapest. This is in accordance with the theoretical worst-case time complexity for each. However, (i) the linked list variants are materially cheaper than the heap variants at dequeuing, and (ii) dequeuing in a call market must be done as a batch, whereas enqueuing is paid for one at a time by the trader submitting the order, and (iii) Ethereum will not permit more than hundreds of orders so the asymptotic behaviour is not a significant factor. For these reasons, we suggest using a linked list variant for this specific application. As it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into a linked list PQ is significantly greater than the linked list with mapping, as each insertion causes the creation of a new DApp. Accordingly, we choose to implement the call market with the linked list with mapping. Overall this PQ balances a moderate gas cost for insertion (\ie order submission) with one for removal (\ie matching the orders).

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =  %

\subsection{Cost/Benefit of Cleaning up After Yourself}
\label{sec:gasrefund}

\input{tables/cleaning.tex}

\paragraph{Gas Refunds.} In order to reconstruct the current state of Ethereum's EVM, a node must obtain a copy of every variable change since the genesis block (or a more recent `checkpoint' that is universally agreed to). For this reason, stored variables persist for a long time and, at first glance, it seems pointless to free up variable storage (and unclear what `free up' even means). Once the current state of the EVM is established by a node, it can forget about every historical variable changes and only concern itself with the variables that have non-zero value (as a byte-string for non-integers) in the current state (uninitialized variables in Ethereum have the value zero by default). Therefore, freeing up variables will reduce the amount of state Ethereum nodes need to maintain going forward.

For this reason, some EVM operations cost a negative amount of gas. That is, the gas is refunded to the sender at the end of the transaction, however (1) the refund is capped at 50\% of the total gas cost of the transaction, and (2) the block gas limit applies to the pre-refunded amount (\ie a transaction receiving a full refund can cost up to 5.5M gas with a 11M limit). Negative gas operations include:

\begin{itemize}

\item \texttt{SELFDESTRUCT}. This operation destroys the contract that calls it and refunds its balance (if any) to a designated receiver address. \texttt{SELFDESTRUCT} operation does not remove the initial byte code of the contract from the chain. It always refunds 24,000 gas. For example, if a contract A stores a single non-zero integer and contract B stores 100 non-zero integers, the \texttt{SELFDESTRUCT} refund for both is the same (24,000 gas).

\item \texttt{SSTORE}. This operation loads a storage slot with a value. Using \texttt{SSTORE} to load a zero into a storage slot means the nodes can start ignoring it (recall that all variables, even if uninitialized, have zero by default). Doing this refunds 15,000 gas per slot.

\end{itemize}

At the time of this writing, Ethereum transaction receipts only account for the \texttt{gasUsed}, which is the total amount of gas units spent during a transaction, and users are not able to obtain the value of the EVM's refund counter from inside the EVM~\cite{signer2018gas}. So in order to account for refunds in Tables~\ref{tab:PQUnitTests}, we calculate them manually. First we need to figure out exactly how much storage is being cleared or how many smart contracts are being destroyed, then we multiply these numbers by 24,000 and 15,000 respectively.

\paragraph{Refunds and PQs.}

One consequence of a linked list is that a new DApp is created for every node in the list. Beyond being expensive for adding new nodes (a cost that will be bared by the trader in a call market), it also leaves a large footprint in the active Ethereum state, especially if we leave the nodes on the blockchain in perpetuity (\ie we just update the head node of the list and leave the previous head `dangling'). However in a PQ, nodes are only removed from the head of the list; thus the node contracts could be `destroyed' one by one using an extra operation, \texttt{SELFDESTRUCT}, in the \texttt{Dequeue()} function. As shown in Table~\ref{tab:cleaning}, the refund from doing this outweighs to the cost of the extra computation: gas costs are reduced from 721K to 557K.  This suggests a general principle: cleaning up after yourself will pay for itself in gas refunds. Unfortunately, this is not universally true as shown by applying the same principle to the linked list with mapping.

Dequeuing in a linked list with mapping can be implemented in two ways. The simplest approach is to process a node, update the head pointer, and leave the `removed' node's data behind in the mapping untouched (where it will never be referenced again). Alternatively, we can call \texttt{DELETE} on each mapping entry once we are done processing a node in the PQ. As it can be seen in the last two rows of Table~\ref{tab:cleaning}, leaving the data on chain is cheaper than cleaning it up.

The lesson here is that gas refunds incentivize developers to clean up storage variables they will not use again, but it is highly contextual as to whether it will pay for itself. Further the cap on the maximum refund means that refunds are not fully received for large cleanup operations (however removing the cap impacts the miners' incentives to include the transaction). This is a complicated and under-explored area of Ethereum in the research literature. For our own work, we strive to be good citizens of Ethereum and clean up to the extent that we can---thus all PQs in Table~\ref{tab:PQUnitTests} implement some cleanup.

% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =
% = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = = =

\section{Call Market Design}

\input{tables/functions.tex}

\cm is proof of concept code. We try to simplify the design at every step to make it an extensible base class but still functional without any extensions. A call market will open for a specified period of time during which it will accept a capped number of orders (\eg 100 orders---parameterized so that all orders can be processed), and these orders are added to a PQ. Our vision (discussed below) is the market would be open for a very short period of time, close, and then reopen immediately (\eg every other block). We keep the design simple by not allowing cancellations which require a second transaction and front-running attacks apply to cancellation orders~\cite{eskandari2019sok}. As markets are relatively short-lived, orders simply expire when the market call period ends.

Another simplifying assumption is to implement a \textit{collateralized} call market. We assume all trades are between ETH and an ERC20 token, all orders are pre-funded in the DApp with ETH (for bids) and tokens (for asks), and once ETH or tokens are committed to an order, they cannot be withdrawn until the market closes. This ensures all executed orders clear and settle (\ie no defaults on payment or delivery).

% Volume of 1
% Ties?

\cm is open source and written in 336 lines (SLOC) of Solidity plus the priority queue (\eg a heap with dynamic array is 282 SLOC). Refer to the availability section at the end of the paper. We tested it with the Mocha testing framework using Truffle on Ganache-CLI to obtain our performance metrics. Once deployed, the bytecode of \cm is 10,812 bytes plus the constructor code (6,400 bytes) which is not stored. We cross-checked for vulnerabilities with \textit{Slither}\footnote{https://github.com/crytic/slither} and \textit{SmartCheck}\footnote{https://tool.smartdec.net} (it only fails some `informational' warnings that are intentional design choices---\eg a costly loop). Table~\ref{tab:cm_functions} summarizes \cm's primary operations.

%============= Performance =================== %

 \subsection{Measurements}

 \input{tables/performance.tex}

The main research question is how many orders can be processed in a single Ethereum transaction when closing the call market, using Ethereum today. As our previous experiments indicated, the choice of PQ implementation is the main influence on performance (see Table~\ref{tab:worst_case_matching}). We implemented a generic call market that interfaces with a generic PQ (at its own contract address) and ran experiments for each PQ implementation. We looked at the \textit{worst-case} for performance which is when every submitted bid and ask is marketable (\ie will require fulfillment). In the first two columns, we determine the highest number of orders that can be processed in a single call to the \texttt{closeMarket()} and not exceed the the current Ethereum block gas limit of 11,741,495. Since Ethereum will become more efficient over time, we also were interested in how much gas it would cost to execute 1000 pairs of orders which is given in the third column. The fourth column indicates the cost of submitting a bid or ask --- since this costs will vary depending on how many orders are already submitted (recall Figure~\ref{fig:random_insertion}), we average the cost of 200 order submissions.

As expected, the numbers closely track the performance of the PQ itself suggesting the PQ is indeed the main influence on performance. In Ethereum today, call markets appear to be limited to processing about a hundred orders per transaction. If markets open on every other block and the call market could monopolize an entire block to close, a few hundred orders per minute (worst-case) can be processed. The main takeaway is that the transparency, front-running resistance, and low barrier to entry of Ethereum come with an enormous cost (\ie an institutional exchange like NASDAQ can process 100K trades per second). That said, many exchanges trade the same assets under different trading rules (\ie market fragmentation) because traders have different preferences. \cm can work today in some circumstances like very low liquidity tokens, or markets with high volumes and a small number of traders (\eg liquidation auctions).

\subsection{Front-running} \label{sec:front}

The primary motivation for using a call market, as opposed to a continuous-time order book, is to mitigate front-running attacks~\cite{clark2014decentralizing,eskandari2019sok,daian2019flash}. Consider a sequence of 100 orders submitted within a window of time to market A, and the same sequence is randomly shuffled and submitted to market B. If A and B are continuous, the orders that are executed in A could be quite different from B. In a call market, the outcome of A and B will be exactly equivalent. There is no threat from miners reordering transactions or traders offering higher gas rates to have their orders executed before other orders already broadcasted.

In a traditional call market, a market clearing price is chosen and all trades are executed at this price. All bids made at a higher price will receive the assets for the lower clearing price (and conversely for lower ask prices): this is called a \textit{price improvement}. A miner about to mine on a set of transactions, including its own orders, could drop other traders' orders to maximize its own price improvement. For this reason, in \cm, all price improvements are given to the miner (using \texttt{block.coinbase.transfer()}). This does not actively hurt traders---they always receive the same price that they quote in their orders---and it removes any incentive for miners to front-run these profits.

Potential for front-running still exists around ties on price and order cancellations (both discussed in the next section). Finally, \textit{displacement attacks}~\cite{eskandari2019sok} are possible where competitive orders are delayed long enough for the market to close without them or the adversary fills up the current market (which caps the number of orders) to ensure other orders cannot be added. Both are expensive attacks. For all these reasons, we say \cm reduces front-running attacks but stop short of saying this issue is completely solved by \cm.

\subsection{Cleaning Up Revisted}

@@

% = = = = = = = = = = = = = = = = = = = = = = = = =  %

\section{Design Alternatives and Extensions}

%\textblue{ We need to find a home for this: another solution to reopen the market immediately after the closing call is to have many similar DApps for the market and  a factory DApp to guide users to the correct address (from the reviewer's comment)}

\cm is designed as base class that can be extended and customized. We discuss potential modifications here.

\subsection{Token Divisibility and Ties}

When executing trades, if the volume of the current best bid does not match the best ask, the larger order is partially filled and the remaining volume is considered against the next best order. In \cm, tokens are assumed to be divisible. We simplify the market rules around ties on price: we execute them in a FIFO manner (breaking front-running resistance for ties on prices). A common trading rule (that does resist front-running) is to fill ties in proportion to their volume (\ie \textit{pro rata} allocation)\footnote{If Alice and Bob bid the same price for 100 tokens and 20 tokens respectively, and there are only 60 tokens left in marketable asks, Alice receives 50 and Bob 10.} however this approach does not always work. Consider the following corner-case: 3 equally priced bids of 1 non-divisible token and 1 ask at the same price. There is no good option: (1) the bids could all be dropped (fair but not market efficient), (2) bids could be prioritized based on time as in \cm (front-running is viable, but in this corner case only), or (3) the bid could be randomly chosen (\cf Libra~\cite{mavroudis2019libra}); blockchain `randomness' is generally deterministic and manipulatable by miners~\cite{bonneau2015random,buenz2017proofs} and counter-measures could take a few blocks to select~\cite{boneh2018verifiable}).

\subsection{Order Cancellations}

\cm does not support order cancellations. We intend to open and close markets quickly (on the order of blocks), so orders are relatively short-lived. Support for cancellation also opens the market to new front-running issues where other traders (or miners) can displace cancellations until after the market closes (however one benefit of a call market is that beating a cancellation with a new order has no effect, assuming the cancellation is run any time before the market closes). Finally, cancellations have a performance impact. Cancelled orders can be removed from the underlying data structure   or accumulated in a list that is cross-checked when closing the market. Removing orders requires a more verbose structure than a priority queue (\eg a binary search tree instead of a heap; or methods to traverse a linked list rather than only pulling from the head). While client software could help point out where the order is in the data structure, the order book can change between submitting the cancellation request and running the method. A linked list with mapping that returns the key for each submitted order seems to be the most tenable data structure.

\subsection{Market Clearing Price}

Call markets are heralded for fair price discovery. This is why many exchanges use a call market at the end of the day to determine the closing price of an asset, which is an important price both optically (it is well-published) and operationally (many derivates settle based on the closing price). We purposely do not compute a `market clearing price' with \cm because miners can easily manipulate the price (\ie include a single wash trade at the price they want fixed), although they forgo profit for doing so. This is not merely hypothetical---Uniswap (the prominent quote-drive, on-chain exchange) prices have been manipulated to exploit other DeFi applications relying on them. Countermeasures to protect Uniswap price integrity could also apply to \cm: (1) taking a rolling mediant of prices over time, and (2) using it alongside other sources for the same price and forming a consensus. While \cm does not emit a market clearing price, it can be computed by a web application examining the order book at market close.

\subsection{Scheduling Events}

As a simple base class, \cm is a one-shot market. However it can be extended to re-open with a clean order book after \texttt{closeMarket()} is run. Modifiers can enforce when the market operates openly (collecting orders) and when close can be run. In the Princeton paper~\cite{clark2014decentralizing}, the call market is envisioned to run as an alt-coin, where orders accumulate within a block and a miner closes the market as part of the logic of producing a new block (\ie within the same portion of code as computing their coinbase transaction in Bitcoin or gasUsed in Ethereum).

If the call market runs as a DApp on Ethereum, it seems difficult to open and close the market every block. Someone needs to call the \texttt{closeMarket()} for every block (we return to who this is next) but the market will only work as intended if miners execute this function after every \texttt{submitBid()} and \texttt{submitAsk()} invocation. Since price improvements are paid to the miners, the miner is actually incentivized to run \texttt{closeMarket()} last to make the most profit. This pattern is called miner extractable value (MEV)~\cite{daian2019flash} and is usually considered in the context of attacks. However in our case, MEV is a feature. Efficient algorithms for miners to automatically find MEV opportunities is an open research problem.

A close alternative is to allow markets to open and close on different blocks. In this alternative, the \texttt{closeMarket()} function calls \texttt{openMarket()} as a subroutine and sets two modifiers: orders are only accepted in the block immediately after the current block (\ie the block that executes the \texttt{closeMarket()}) and \texttt{closeMarket()} cannot be run again until two blocks after the current block.

>>>>

\section{Availability} 
The Solidity source code for \cm, Truffle test files, and Arbitrum dependencies are available in a GitHub repository (non-anonymous link omitted). We have also deployed \cm on Ethereum's testnet Rinkeby with flattened (single file) source code of just the \cm base class and priority queue implementations. It is visible and can be interacted with here: \href{https://rinkeby.etherscan.io/address/0x0d91de29c531d074853a5cef7cf9dfeb9c6ec4e0}{[etherscan.io]}; while the Arbitrum variant is here: \href{https://explorer5.arbitrum.io/#/address/0x18BF65069Fdaed79f8BDB0B6EDa87d5cc46F6E28}{[explorer5.offchainlabs.com]}.

\section{Roll-ups: Additional Background}
\label{app:rollup}

\subsection{Cleaning Up Revisited}
\label{app:clean}

@@
